Nail: A practical interface generator for data formats

### 2014-05-01 paper/osdi14/impl.tex
Implementation

The current prototype of the Nail parser generator supports the C programming
language and top-down parsers. The implementation parses Nail grammars with Nail itself, using a
130-line Nail grammar feeding into a 2,000-line C++ program that
implements the parser. Bootstrapping is
performed with a subset of the grammar implemented using conventional grammars. 
Options for C++ STL data models and emitting
Packrat parsers~\cite{packrat-parsing:icfp02} are under development. In
this section, we will discuss some particular features of our parser
implementation.

@@

A generated Nail parser makes two passes through the input: the first to
validate and recognize the input, and the second to bind this data to the internal
model. Currently the parser is a straightforward top-down parser, although
preparations have been made to add Packrat parsing to achieve linear parsing
times.

\paragraph{Memory allocation.}

Many security vulnerabilities can occur when heap allocations are done improperly. Therefore, just
like Hammer, Nail avoids using the heap as much as possible, using a custom arena allocator and
allocating only fixed-size blocks from the system allocator (malloc). However, Nail extends upon
Hammer's approach and uses two arenas for each parsed input. One arena is used for intermediate
structures, including dependency fields and any temporary streams, and is released (and zeroed)
after parsing completes, whereas the other arena is used only for allocating the result, and has to
be freed by the programmer.

\paragraph{Intermediate representation.}

Most parser generators, such as Bison, do not have to dynamically allocate temporary data on the
heap, as they evaluate a semantic action on every rule. However, as our goal is to perform as little
computation as possible before the input has been validated, and we do not want to mix temporary
objects with the results of our parse, we use an append-only trace to store intermediate parser
results.

Hammer solves this problem by storing a full abstract syntax tree. However, this
abstract syntax tree is at least an order of magnitude larger than the input,
because it stores a large tree node structure for each input byte and for each
rule reduced. This allows Hammer semantic actions to get all of the necessary
information without ever seeing the raw input stream. However, because we also
automatically generate our second pass, which corresponds to Hammer's semantic
actions, we can trust it as much as we trust the parser, and thus can expose it
to the raw input stream.

Under this premise, the actions need limited information from the
recognizer to correctly handle the input stream. In particular, the parser's
control flow branches only at the choice, repetition, and constant combinators.
Thus, for each of those combinators, we store the minimum amount of information
required to reconstruct the syntactic structure of the input.

The trace is an
array of integers.
Whenever the parser encounters a choice, it appends two integers to the trace:
the number of that choice and
the length of the trace when it began parsing that choice. When backtracking in
the input, the parser does not backtrack in the trace. This means that offsets
within the trace can be used for a Packrat hash table to memoize backtrack-heavy
parsers.
When encountering a repetition combinator, the parser records the number of
times the inner parser succeeded, and when encountering a constant parser of
variable size, it records how much input was consumed by the constant parser.

In a second pass, the parser then allocates the internal representation from an
arena allocator and binds the fields to values from the input, while following
the trace to determine how many array fields to parse and which choices to pick.

>>>>

The current prototype of the Nail parser generator supports the C programming
language. The implementation parses Nail grammars with Nail itself, using a
130-line Nail grammar feeding into a 2,000-line C++ program that
emits the parser and generator code. Bootstrapping is
performed with a subset of the grammar implemented using conventional grammars. 
And option for C++ STL data models  is in development. In
this section, we will discuss some particular features of our parser
implementation.

@@

A generated Nail parser makes two passes through the input: the first to
validate and recognize the input, and the second to bind this data to the internal
model. Currently the parser uses a straightforward top-down algorithm, which can perform poorly on
grammars that backtrack heavily. However, preparations have been made to add Packrat
parsing~\cite{packrat-parsing:icfp02} that achieve linear time even in the worst case.

\paragraph{Defense-in-Depth.}

Security exploits often rely on raw inputs being present in memory, for example to include
shell-code or crafted stack frames for ROP-attacks in padding fields\cite{shotgun-parser} or the
application executing a controlled sequence of heap allocations and de-allocations to place
specific data at predictable addresses\cite{jp-advanced, vudo-malloc}. Because the rest of the
application or even Nail's generated code may contain memory corruption bug, Nail
carefully handles memory allocations as defense-in-depth to make exploiting such vulnerabilities harder.

When parsing input, Nail uses two separate memory arenas. These arenas allocate memory from the
system allocator in large, fixed size blocks. Allocations are handled linearly and all data in the
arena is zeroed and freed at the same time. Nail uses one arena for data used only during parsing,
including dependent fields and temporary streams and is released before the Nail's parser returns.
The other arena is used to allocate the internal data type returned and is freed by the application
once it is done processing an input.

Furthermore, the internal representation does not include any references to the input stream, which
can therefore be zeroed immediately after the parser succeeds, so an attacker has to write an
exploit that works without referencing data in the raw input.
% \paragraph{Intermediate representation.}

% Most parser generators, such as Bison, do not have to dynamically allocate temporary data on the
% heap, as they evaluate a semantic action on every rule. However, as our goal is to perform as little
% computation as possible before the input has been validated, and we do not want to mix temporary
% objects with the results of our parse, we use an append-only trace to store intermediate parser
% results.

% Hammer solves this problem by storing a full abstract syntax tree. However, this
% abstract syntax tree is at least an order of magnitude larger than the input,
% because it stores a large tree node structure for each input byte and for each
% rule reduced. This allows Hammer semantic actions to get all of the necessary
% information without ever seeing the raw input stream. However, because we also
% automatically generate our second pass, which corresponds to Hammer's semantic
% actions, we can trust it as much as we trust the parser, and thus can expose it
% to the raw input stream.

% Under this premise, the actions need limited information from the
% recognizer to correctly handle the input stream. In particular, the parser's
% control flow branches only at the choice, repetition, and constant combinators.
% Thus, for each of those combinators, we store the minimum amount of information
% required to reconstruct the syntactic structure of the input.

% The trace is an
% array of integers.
% Whenever the parser encounters a choice, it appends two integers to the trace:
% the number of that choice and
% the length of the trace when it began parsing that choice. When backtracking in
% the input, the parser does not backtrack in the trace. This means that offsets
% within the trace can be used for a Packrat hash table to memoize backtrack-heavy
% parsers.
% When encountering a repetition combinator, the parser records the number of
% times the inner parser succeeded, and when encountering a constant parser of
% variable size, it records how much input was consumed by the constant parser.

% In a second pass, the parser then allocates the internal representation from an
% arena allocator and binds the fields to values from the input, while following
% the trace to determine how many array fields to parse and which choices to pick.
